Improving the communication and transparency of stock assessment using interactive

### 2019-01-14 analysis/paper/manuscript.Rmd
Draft section on NCAM dashboard

# Introduction

@@

## Knowing the data

@@

- *Greg, can you introduce this tool, providing some info on the tagging database and the tools used to make the tool (Shiny, leaflet)?*
- *TODO: the intro paragraph to this section could use some references if it is necessary*

## Data modeling

@@

The Northern cod stock off southern Labrador and eastern Newfoundland is one of the most well studied stocks in eastern Canada, perhaps by virtue of its history. As such, there are multiple monitoring programs that help inform the status of the stock and data from most of these programs have been integrated into a state-space stock assessment model, called NCAM [@cadigan2016]. The "base case" model includes information from research vessel autumn trawl surveys (1983-present), Sentinel fishery surveys (1995-present), inshore acoustic surveys (1995-2009), fishery catch-at-age compositions and partial fishery landings (1983-present), and tagging (1983-present). Using a series of observation equations, this TMB based model reduces thousands of historical data points from these monitoring programs into quantities such as recruitment, spawning stock biomass, fishing mortality and natural mortality. Once the model is fit to the data, the next step is to produce visual representations of the data and model. An interactive and self-contained "dashboard" was developed for this purpose for the last assessment of Northern cod [@dwyer2019]. The concept of using a dashboard was burrowed from the business community where dashboards are frequently used to group a series of interactive visuals and tables to provide at-a-glance views of key performance indicators. Many tools have been developed to facilitate the development of data dashboards and we leveraged R-based tools to construct a tool for exploring the input and output of NCAM. Specifically we utilized the flexdashboard [@allaire2017] package to group interactive plotly-based [@plotly] visuals into a dynamic document. We also used the crosstalk [@cheng2016] package to link the data displayed across multiple plots in the "Tagging" tab.

TODO:
- Take screenshots of the dashboard for static figures for the paper (use webshot?)
- Remember to add reference to the supplement that will ultimately include the NCAM_dashboard.html

# Notes

- Note that these tools facilitate the peer-review process and also improves the transparency and reproducibility of assessment models
- Dynamic document can be self contained
- Northern cod stock is on an annual assessment schedule, therefore it makes sense to automate visuals

# Other draft text

A wide array data may be collected for an assessment and, formally, the process often reduces to algorithms that convert these data to advice for fisheries managers. In some cases, particularly for commercially valuable species, this means that hundreds or thousands of historical data points from the monitoring program of a stock gets reduced into a single policy value, such as a recommended catch quota [@maunder2009]. Sound management advice therefore hinges on the ability of stock assessment biologists to manage, interpret and communicate large volumes of data from a variety of sources.

In this data-intensive era of stock assessment, one of the difficulties becomes visualizing the inputs and outputs to integrated stock assessment models. The standard approach has been to create a series of tables and static plots to help assess the inputs and model fits. Data presented in tables is incredibly valuable, but as human beings, seeing patterns can be challenging. Static figures, in contrast, help reveal patterns but they are divorced from the underlying data and, as such, limit detailed explorations. New tools are bridging this gap by making it easier to build interactive figures [@perkel2018].

Interactive visuals...visual representation of the data and model that help us interpret complex data at a glance https://campus.sagepub.com/blog/3-benefits-of-interactive-visualization

https://centricdigital.com/blog/digital-trends/the-value-of-interactive-data-visualization-and-why-it-matters-to-business/

https://www.nature.com/articles/s41467-018-03297-7

https://esajournals.onlinelibrary.wiley.com/doi/full/10.1890/120103

https://www.nature.com/articles/d41586-018-01322-9

"The ultimate objective of Fisheries Science is to inform management" (http://www.fao.org/tempref/docrep/fao/008/a0212e/a0212E02.pdf)

>>>>

@@

# Knowing the data

@@

> Greg, can you introduce this tool, providing some info on the tagging database and the tools used to make the tool (Shiny, leaflet)?

> TODO: the intro paragraph to this section could use some references if it is necessary

# Data modeling

@@

The Northern cod stock off southern Labrador and eastern Newfoundland is one of the most well studied stocks in eastern Canada, perhaps by virtue of its history. As such, there are multiple monitoring programs that help inform the status of the stock and data from most of these programs have been integrated into a state-space stock assessment model, called NCAM [@cadigan2016]. The "base case" model includes information from research vessel autumn trawl surveys (1983-present), Sentinel fishery surveys (1995-present), inshore acoustic surveys (1995-2009), fishery catch-at-age compositions and partial fishery landings (1983-present), and tagging (1983-present). Using a series of observation equations, this TMB based model reduces thousands of historical data points from these monitoring programs into quantities such as recruitment, spawning stock biomass, fishing mortality and natural mortality. Once the model is fit to the data, the next step is to produce visual representations of the data and model. An interactive and self-contained "dashboard" was developed for this purpose for the last assessment of Northern cod [@dwyer2019]. The concept of using a dashboard was burrowed from the business community where dashboards are frequently used to group a series of interactive visuals and tables to provide at-a-glance views of key performance indicators. Many tools have been developed to facilitate the development of data dashboards and we leveraged R-based tools to construct a tool for exploring the input and output of NCAM. Specifically we utilized the flexdashboard [@allaire2017] package to group interactive plotly-based [@plotly] visuals into a dynamic document. We also used the crosstalk [@cheng2016] package to link the data displayed across multiple plots.

The NCAM dashboard, which is included in Supplement 2 as a self-contained html file, contains a series of tabs, the first of which provides terse point-form background on the model (tab named "Background"). Subsequent tabs provide a series of diagnostic plots for assessing model fits to the catch ("Catch"), research vessel autumn trawl survey ("RV survey"), Sentinel fishery survey ("SN survey"), inshore acoustic survey ("SS survey"), and tagging ("Tagging") data. For instance, the "RV survey" tab includes plots of observed and predicted values of mean numbers per tow captured in the research vessel survey (Figure \@ref(fig:rv)). The dashboard also includes tabs focused on model estimates such as catchability and selectivity ("Catchability"), stock size and vital rates ("Trends"; Figure \@ref(fig:trends)), and stock productivity ("Productivity"). Finally, some results from a retrospective analyses are included under the "Retro" tab and key inputs and outputs are included as tables under the "Tables" tab. The plots and tables included in the dashboard are similar to those typically presented at assessment meetings and in research documents, however, the interactive nature of the dashboard permits easy and efficient access to significantly more detail. Because the dashboard is contained in an html file, it acts as an interactive assessment document that can be shared for peer-review. This allows colleagues and stakeholders to independently scrutinize details of the data and model that are not easily accessed by users other than the analyst. Such access improves the transparancy of the stock assessment model which, in turn, leads to richer discussion and interogation of the biological and statistical rigor of the model.

> TODO: should I add detail of questions not previously raised? If so, I can't quite remember.

```{r rv, echo = FALSE, fig.cap = '- Screenshot of the "RV survey" tab from the NCAM dashboard where total observed (dots) and NCAM model predicted values (lines) for the DFO RV survey index are shown in the upper left panel and scaled matrix plot of age-disaggregated standardized log residuals are shown in the lower left panel (blue = positive, red = negative, symbols scaled by size; grey = index values of zero). Age-disaggregated observed (dots) and perdicted (lines) values from the DFO RV survey are shown in the right panel. The tab on the right panel including residual plots are hidden here, but asscessable in the dashboard.'}
knitr::include_graphics("figures/NCAM_dashboard_rv-survey_tab.png")
```

```{r trends, echo = FALSE, fig.cap = '- Screenshot of the "Trends" tab which displays estimates of recruitment, stock size, and stock size relative to B~lim~ (left panels) and mortality rates (F, M, Z, right panels).'}
knitr::include_graphics("figures/NCAM_dashboard_trends_tab.png")
```

# Open stock-assessment

> TODO: discuss this concept in the context of open science and open stock assessment.

> **Other draft text**

> A wide array data may be collected for an assessment and, formally, the process often reduces to algorithms that convert these data to advice for fisheries managers. In some cases, particularly for commercially valuable species, this means that hundreds or thousands of historical data points from the monitoring program of a stock gets reduced into a single policy value, such as a recommended catch quota [@maunder2009]. Sound management advice therefore hinges on the ability of stock assessment biologists to manage, interpret and communicate large volumes of data from a variety of sources.

> In this data-intensive era of stock assessment, one of the difficulties becomes visualizing the inputs and outputs to integrated stock assessment models. The standard approach has been to create a series of tables and static plots to help assess the inputs and model fits. Data presented in tables is incredibly valuable, but as human beings, seeing patterns can be challenging. Static figures, in contrast, help reveal patterns but they are divorced from the underlying data and, as such, limit detailed explorations. New tools are bridging this gap by making it easier to build interactive figures [@perkel2018].

> **Links to some handy references**

> Interactive visuals...visual representation of the data and model that help us interpret complex data at a glance https://campus.sagepub.com/blog/3-benefits-of-interactive-visualization

> https://centricdigital.com/blog/digital-trends/the-value-of-interactive-data-visualization-and-why-it-matters-to-business/

> https://www.nature.com/articles/s41467-018-03297-7

> https://esajournals.onlinelibrary.wiley.com/doi/full/10.1890/120103

> https://www.nature.com/articles/d41586-018-01322-9

> "The ultimate objective of Fisheries Science is to inform management" (http://www.fao.org/tempref/docrep/fao/008/a0212e/a0212E02.pdf)
